A Machine Profile Summary

### 2022-01-18 MachineSummary/MachineSummary.Rmd
update MachineSummary

---
title: "A Machine Profile Summary"
author: 
  - John C. Nash, University of Ottawa, Canada
date: "14/06/2021"
bibliography: ../BibSupport/ImproveNLS.bib
output: 
  pdf_document:
    keep_tex: false
---

@@

I have looked at this.

If we want to compare results computed on different machines, we need a way to provide 
measures and identifiers of the particular computing environment at hand. This document
attempts to summarize some possibilities available in the R statistical software and language.

@@

There are multiple ways to get information of the type useful for characterizing the
computing infrastructure used for a particular test or timing. We will need to
ensure that identifiers of particular pieces of information are unique, and that 
common identifiers really do specify identical information. Otherwise, we need
to develop new names for our set of information.

@@

Most computatins should have a time/date stamp that is unlikely to be confused with any
other, at least in combination with a machine name and/or other tags.

```{r tdstamp}
#get a timestamp
tsstr<- format(Sys.time(), "%Y%m%d%H%M") # tsstr == time stamp string in form YYYYmmddHHMM
cat("Date and Time stamp:",tsstr,"\n")
```
## Identifier for the computing environment

@@

- machine name. While most systems (and Linux in particular) offer to let the user
  provide a machine name, there are generally defaults that many people accept, and
  these are often uninformative and may be non-unique. We note that VirtualBox with
  a Windows 10 guest machine gave the name DESKTOP-HF4CKVA. Where this name was 
  generated we are not sure. Settings / System allows "rename this PC".

- operating system and version

@@

- cpu
- cores
- operating cycle speed (While the cpu is generally specified )
- RAM total
- RAM available, possibly indicating swap space and usage
- RAM speed (could be variable)

@@

?? what to put in.

## R tools for machine information

### sessionInfo()

@@

### benchmarkme

@@

### calceps.R

This is a port of Mike Malcolm's ENVRON to R. ??ref. It computes the 
machine precision, radix and number of radix digits.

?? get file into chunk

```
calceps.R

@@

### Sys.info()

@@

## Issues relateing to RNG

@@

```
sys.frame()
sys.frames()
sys.status()
sys.on.exit()
sys.parents()
sys.calls()
```

# Choices

>>>>

---
title: "A Machine Profile Summary"
author: 
   - John C. Nash \thanks{ retired professor, Telfer School of Management, University of Ottawa}
   - Arkajyoti Bhattacharjee \thanks{Department of Mathematics and Statistics, Indian Institute of Technology, Kanpur}
date: "14/06/2021 revised 18/01/2022"
bibliography: ../BibSupport/ImproveNLS.bib
output: 
  pdf_document:
    keep_tex: false
---

@@

If we want to compare results computed on different machines, we need a way to provide 
measures and identifiers of the particular computing environment at hand. This write-up
attempts to summarize some possibilities available in the R statistical software and language.

@@

There are multiple ways to get information of the type useful for characterizing the
computing infrastructure used for a particular test or timing. However, we need to
ensure that identifiers of particular pieces of information are unique, and that 
common identifiers really do specify identical, or at least equivalent, information. 
Otherwise, new names for some pieces of our set of capability information are needed.

@@

When we wish to analyze performance results, it is extremely helpful to have 
information on when the calculations were performed. 
Most computations should have a time/date stamp that is unlikely to be confused with any
other, at least in combination with a machine name and/or other tags. Here is one
suggestion based on the R `Sys.time()` function.

```{r tdstamp}
#get a timestamp
tsstr<- format(Sys.time(), "%Y%m%d%H%M") # tsstr == time stamp string in form YYYYmmddHHMM
cat("Date and Time stamp:",tsstr,"\n")
```

## Identifier for the computing software environment

@@

- a machine name. While most systems (and Linux in particular) offer to let the user
  provide a machine name, there are generally defaults that many people accept, and
  these are often uninformative and may be non-unique. We note that VirtualBox with
  a Windows 10 guest machine gave the name DESKTOP-HF4CKVA. Where this name was 
  generated we are not sure. Settings / System allows "rename this PC".

- operating system and version. For Linux systems, we note that besides the distribution
  variant, e.g., Linux Mint 20.3 MATE 64 bit, it is necessary to add the Linux kernel
  identifier, e.g.,  5.13.0-25-generic, from the `uname` command.

@@

- cpu (i.e., processor)
- number of cores
- operating cycle speed (this is sometimes specific to the cpu specified )
- RAM size total
- RAM size available, possibly indicating swap space and usage
- RAM speed (which could be variable)
- GPU (i.e, additional processors) information.

@@

Clearly it is important to identify which software packages are active, and the
R `sessionInfo()` command is useful for this. However, there may be additional 
features that should be mentioned if the user has modified the software in any
way.

The use of tools to carry out computations in parallel, using multiple cores or
GPUs, is clearly an issue. In our opinion, such possibilities are not well supported
by common R functions, or indeed in other computing languages, though we would very
much like to hear of tools to provide such information.

# R tools for machine information

## sessionInfo()

@@

## benchmarkme

?? Need to comment on

@@

## calceps.R

Until the IEEE 754 standard (??ref) became availaible in the 1980s, 
computers used multiple forms of floating point arithmetic. With
some ingenuity, it is possible to detect many aspects of the 
floating point properties.
This is a port of Mike Malcolm's ENVRON to R. ??ref. It computes the 
machine precision, radix and number of radix digits.

?? mention Kahan's Paranoia

```{r envrontest}
envron<-function(){
D1<-as.numeric(1) # use variable to avoid constants when double precision invoked
E5 <- 10 #arbitrary scaling for additive equality tests
B9 <- 1E+35 # big number, not necessarily biggest possible
E6 <- 1 # initial value for radix 
E9 <- 1 # initial value for machine precision
D0 <- E6 
repeat{
  E9 <- E9/2 #start of loop to decrease estimated machine precision
  D0 <- E6+E9 #force storage of sum into a floating-point scalar
  if (D0 <= E6) break
} # repeat reduction while (1+E9) > 1
E9 <- E9*2 # restore smallest E9 which gives (1+E9) > 1
repeat {
   E6 <- E6+1 # try different radix values
   D0 <- E6+E9
   if (D0 <= E6) break
} # until a shift is observed
J1 <- 1 # initial count of radix digits in mantissa
E9 <- 1 # use radix for exact machine precision
repeat {  
  E9 <- E9/E6 #loop while dividing by radix
  J1 <- J1+1 #increment counter
  D0 <- D1+E9 #add tp 1
  if (D0<=D1) break #test and repeat until equality
}
E9 <- E9*E6 #recover last value of machine precision
J1 <- J1-1 #and adjust the number of digits
mpvals<-list(eps=E9, radix=E6, ndigits=J1)
}
mp<-envron()
E9<-mp$eps
E6<-mp$radix
J1<-mp$ndigits
cat("E9 = ",E9,"  -- the machine precision,  E9=MIN (X  1+X>1)\n")
cat( "E9*1E+16=",E9*1E+16,"\n")
cat( "E6 = ",E6,"  -- the radix of arithmetic","\n")
cat( "J1 = ",J1,"  -- the number of radix digits in mantissa of FP numbers","\n")
cat( "E6^(-J1+1) * 1E+16=",1E+16*(E6^(-J1+1)),"\n")
cat(".Machine$double.eps=",.Machine$double.eps,"\n")
```

@@

## Sys.info()

@@

## Issues relating to random number generation

@@

```
sys.frame()
sys.frames()
sys.status()
sys.on.exit()
sys.parents()
sys.calls()
```
# Choices
